---
title: "ISLR-R: Appendix C. Exercises"
output: html_document
date: "2023-03-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise (Lab 1)

Prerequisites:

```{r}
import::from(magrittr, "%>%", "%$%", .into = "operators")
import::from(tidyr, "replace_na")
# load data
auto <- read.csv("../data/Auto.csv", header = T, na.strings = "?")
```

## Data Imputation

**Replace all missing values in each column with the mean value of that
column.**

```{r cars}
df <- data.frame(
  player = c("A", "B", NA, "D", "E"),
  points = c(17, 13, NA, 9, 25),
  rebounds = c(3, 4, NA, NA, 8),
  blocks = c(1, 1, 2, 4, NA)
)
```

Select numeric / non-numeric column:

```{r}
# numeric
df %>%
  dplyr::select(is.numeric %>% dplyr::where()) %>%
  head()
```

```{r}
# non-numeric
df %>%
  dplyr::select(!is.numeric %>% dplyr::where()) %>%
  head()
```

Replace `NA` with mean (numeric) and empty string (non-numeric) values:

```{r}
df %>%
  dplyr::mutate(dplyr::across(
    is.numeric %>% dplyr::where(),
    ~ replace_na(., mean(., na.rm = T))
  )) %>%
  dplyr::mutate(dplyr::across(
    !is.numeric %>% dplyr::where(),
    ~ replace_na(., "")
  ))
```

## Query

**Select cars with model years between 1970 and 1975 and engine displacement
greater than or equal to 3, and calculate the average fuel efficiency.**

```
SELECT avg(mpg)
  FROM auto
 WHERE year > 1970 AND year <= 1975
   AND displacement >= 3
;
```

Filtering:

```{r}
auto %>%
  dplyr::select("year", "displacement", "mpg") %>%
  dplyr::filter(year %>% dplyr::between(., 70, 75) & displacement >= 3) %>%
  head()
```

```{r}
auto %>%
  dplyr::select("year", "displacement", "mpg") %>%
  dplyr::filter(year %>% dplyr::between(., 70, 75) & displacement >= 3) %>%
  dplyr::summarise(avg_mpg = mean(mpg))
```
## Distribution

Compute the distribution of car brands(names) and the mean weight.

```
SELECT name
     , avg(weight) AS avg_weight
  FROM auto
 GROUP BY 1
;
```

Multiple records with the same name:

```{r}
auto %>%
  dplyr::filter(name == "chevrolet impala") %>%
  head()
```

Summarize:

```{r}
auto %>%
  dplyr::group_by(name) %>%
  # only summarized columns are selected
  dplyr::summarize("count" = dplyr::n(), weight = mean(weight)) %>%
  head()
```

Show in descending order:

```{r}
auto %>%
  dplyr::group_by(name) %>%
  dplyr::summarize("count" = dplyr::n(), weight = mean(weight)) %>%
  dplyr::arrange(dplyr::desc(weight)) %>%
  View
```
