---
title: "ISLR-R: Appendix A. Basic Stats"
output: html_document
date: "2023-03-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
import::from(magrittr, "%>%", "%$%", .into = "operators")
```

## Class vs. Typeof vs. Mode

In R, `mode`, `class`, and `typeof` are functions used to determine the type of
an object.
However, they are not exactly the same thing and are used in different contexts.

`mode` returns six basic modes:

- "logical"
- "numeric"
- "integer"
- "complex"
- "character"
- "raw"
  
It also includes "list" and "expression" as modes, which are not basic modes.
`mode` can be used on any R object.

`typeof` returns the type of object in a lower-level sense than mode.
It returns one of seven types:

- "logical"
- "integer"
- "double"
- "complex"
- "character"
- "raw"
- "list"

`class` is based on R's object-oriented class hierarchy. 

```{r}
x <- 1:5
mode(x)  # "numeric"
class(x) # "integer"
typeof(x) # "integer"

y <- data.frame(a = 1:5, b = letters[1:5])
mode(y) # "list"
class(y) # "data.frame"
typeof(y) # "list"
```

## Dataset and Models

```{r}
names(cars) # column names
row.names(cars) # row names
summary(cars)
str(cars) # also summary
head(cars) # equivalent to cars[1:6,]
tail(cars)
attributes(cars)
# formula
cars %$% class(dist ~ speed) # formula: lhs response; rhs predictor
class(a ~ b)
plot(dist ~ speed, cars)
cars %$% plot(cars$speed, cars$dist)
```

## Statistics

```{r}
x <- round(runif(20, 0, 20), digits = 2)
x
summary(x)
min(x)
max(x)
median(x)
mean(x)

var(x)
sd(x)
sqrt(var(x))
rank(c(2, 4, 6, 0, 1, 4))
rank(c(2, 4, 5, 0, 1, 4), ties.method = "last")
order(x)
order(x, decreasing = T)
sort(x)
x[order(x)]
sort(x, decreasing = T)
x[order(x, decreasing = T)]

sum(x)
length(x)
round(x, 1)

fivenum(x)
quantile(x, c(0, 0.25, 0.5, 0.75, 1))
quantile(x)
```

## TBD

```{r}
mad(x) # "median average distance":

cummax(x) # cumulative maximum
cummin(x) # cumulative minimum
cumprod(x) # cumulative product
cor(x, sin(x / 20)) # correlation coefficient
```
