---
title: "ISLR-R: 2. Regression Models"
output: html_document
date: "2023-02-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
import::from(magrittr, "%>%", "%$%", .into = "operators")
```

In this lab you will work through Section 3.6.
The R code from Section 3.6 is given below.

## Boston Dataset

The name for this dataset is simply "boston".
It has two proto-tasks:

- `nox`, in which the nitrous oxide level is to be predicted.
- `medv`, in which the median value of a home is to be predicted.

Columns:

-   `crim`: per capita crime rate by town
-   `zn`: proportion of residential land zoned for lots over 25,000 sq.ft.
-   `indus`: proportion of non-retail business acres per town.
-   `chas`: Charles River dummy variable (1 if tract bounds river; 0 otherwise)
-   `nox`: nitric oxides concentration (parts per 10 million)
-   `rm`: average number of rooms per dwelling
-   `age`: proportion of owner-occupied units built prior to 1940
-   `dis`: weighted distances to five Boston employment centres
-   `rad`: index of accessibility to radial highways
-   `tax`: full-value property-tax rate per \$10,000
-   `ptratio`: pupil-teacher ratio by town
-   `black`: 1000(Bk: 0.63)\^2 where Bk is the proportion of blacks by town
-   `lstat`: % lower status of the population
-   `medv`: Median value of owner-occupied homes in \$1000's

```{r}
MASS::Boston %>% names()
```

Visual Exploration of `lstat` and `medv`:

```{r}
plot(MASS::Boston$lstat, MASS::Boston$medv)
```

## Simple Linear Regression:

```{r}
mdl <- lm(medv ~ lstat, data = MASS::Boston)
```

There are 12 components inside a fitted model:

```{r}
mdl %>% names()
```

```{r}
mdl %>% coef
```

```{r}
mdl %>% summary
```

```{r}
plot(MASS::Boston$lstat, MASS::Boston$medv, pch = "+")
abline(mdl, lwd = 3, col = "red")
```

### Validation Plots

- Fitted vs Residuals
  - Whether linearity holds.
    This is indicated by the mean residual value for every fitted value region
    being close to 0.
    In R this is indicated by the red line being close to the dashed line.
  - Whether homoskedasticity holds, i.e. the variance is the same.
    The spread of residuals should be approximately the same across the x-axis.
  - Whether there are outliers.
    This is indicated by some ‘extreme’ residuals that are far from the rest.
- QQ Plot
  - https://boostedml.com/2019/03/linear-regression-plots-how-to-read-a-qq-plot.html
  - on the x-axis are the theoretical quantiles of a standard normal
  - For the y-axis, consider the empirical distribution function of the
    standardized residuals

```{r}
par(mfrow = c(2, 2))
plot(mdl)
# qnorm((1 - 0.5) / 506) = -3.09
# qnorm((506 - 0.5) / 506) = 3.09
```

```{r}
plot(predict(mdl), residuals(mdl))
plot(predict(mdl), rstudent(mdl))
plot(hatvalues(mdl))
# which.max(hatvalues(mdl))
```

## Multiple Linear Regression

```{r}
mdl2 <- lm(medv ~ lstat + age, data = MASS::Boston)
mdl2 %>% summary()
mdl3 <- lm(medv ~ ., data = MASS::Boston)
mdl3 %>% summary()
mdl4 <- lm(medv ~ . - age, data = MASS::Boston)
mdl4 %>% summary()
```

## Non-linear Transformations of the Predictors

```{r}
mdl5 <- lm(medv ~ lstat + I(1/lstat), data = MASS::Boston)
mdl5 %>% summary()
par(mfrow=c(2,2))
plot(mdl5)
```

```{r}
mdl6 <- lm(medv ~ I(1/lstat), data = MASS::Boston)
mdl6 %>% summary
```
