---
title: "ISLR-R: 6. Association Analysis"
output:
  prettydoc::html_pretty:
    toc: true
    theme: cayman
    highlight: github
    df_print: paged
date: "2023-05-17"
editor_options:
  markdown:
    wrap: 79
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
import::from(magrittr, "%>%", "%$%", .into = "operators")
```

## Association Analysis using the Apriori algorithm

> need the `arules` package.

We can then read in the .csv file, converting it to a list of transactions:

```{r fig.height=12, fig.width=12}
data <- arules::read.transactions(
  file = "../data/Basket1.txt",
  format = "basket", sep = ","
)
arules::inspect(data)
```

### Frequent Item Sets

We can mine the frequent itemsets, view all the frequent itemsets, and view the
five most frequent itemsets with the greatest support using the following code:

```{r}
items <- arules::apriori(data,
  parameter = list(
    support = 0.4,
    confidence = 0.8, target = "frequent itemsets"
  )
)
arules::inspect(items)
arules::inspect(arules::sort(items, by = "support")[1:5])
# cex.names adjusts the size of the labels on the bars
arules::itemFrequencyPlot(data, support = 0.1, cex.names = 0.8)
```

### Maximal Frequent Item Sets

We can mine the maximal frequent itemsets, and view all the maximal frequent
itemsets using this code:

```{r}
MaxFreqItemsets <- arules::apriori(data,
  parameter = list(
    support = 0.3,
    target = "maximally frequent itemsets"
  )
)
arules::inspect(MaxFreqItemsets)
summary(MaxFreqItemsets)
```

### Closed Frequent Item Sets

We can mine the closed frequent itemsets, and view all the closed frequent
itemsets using this code:

```{r}
ClosedFreqItemsets <- arules::apriori(data,
  parameter = list(
    support = 0.4,
    target = "closed frequent itemsets"
  )
)
summary(ClosedFreqItemsets)
arules::inspect(ClosedFreqItemsets)
```

### Rules

We can mine the association rules, view all the rules, and view the five rules
which have the highest confidence as follows:

```{r}
rules <- arules::apriori(data,
  parameter = list(
    minlen = 2, support = 0.4,
    confidence = 0.8, target = "rules"
  )
)
arules::inspect(rules)
summary(rules)
arules::inspect(arules::sort(rules, by = "confidence")[1:5])
```

Finally, we can find particular rules using this code:

```{r}
rules <- arules::apriori(data,
  parameter = list(
    minlen = 2, support = 0.4,
    confidence = 0.8, target = "rules"
  ),
  appearance = list(default = "lhs", rhs = "Milk")
)
summary(rules)
arules::inspect(arules::sort(rules, by = "confidence"))
```

You can type `?apriori` in the console and then click on 'APappearance' to see
how to extract more complicated rules.

### Interpretation

Parameters:

- additional rules
  - `arem`: additional rule evaluation measure
  - `minval`: its minimal value
  - `aval`: whether to compute additional evaluation measures
- `originalSupport`:  indicating whether or not use the original definition of
  minimum support
  - If set to `TRUE` (which is the default), the support of a rule is
    considered as the frequency of both the LHS and RHS of the rule appearing
    together in the transactions.
  - If set to `FALSE`, the support of a rule is considered as the frequency of
    just the LHS of the rule appearing in the transactions (also known as the
    coverage of the rule).
- `minlen`: minimum number of items that a rule must have.
- `maxlen`: maximum number of items that a rule can have.
- `maxtime`: time limit in seconds for checking subsets.
  `0` disables the time limit. (default: `5`)

Target:

- `frequent itemsets`: frequency of occurrence is defined by the support
  parameter, which is a threshold percentage. (no confidence)
- `rules`: both support and confidence
- `max`: none of its immediate supersets are frequent
- `closed`: none of its immediate supersets have the same support count
- maximally frequent itemsets are the largest sets of items that frequently
  appear together, while closed frequent itemsets are those that don't have any
  immediate supersets with the same support count.

### Formula

$$\rm{Supp} (X \cup Y) = \frac{\sigma(X \cup Y)}{N} = Pr(X, Y)$$
$$\rm{Conf} (X \to Y) = \frac{\sigma(X \cup Y)}{\sigma(X)} = Pr(Y | X)$$
$$\rm{Lift} (X \to Y) =
\frac{\rm{Conf}(X \to Y)}{\rm{Supp} (Y)} = \frac{Pr(X, Y)}{Pr(X)Pr(Y)}$$

## Other data sets to play with for association analysis

These are much larger, real-world data sets.

```{r}
# 1. Adult Census Income Database
data("Adult", package = "arules")
Adult %>% head()
# 2. Groceries data set contains 1 month (30 days) of real-world point-of-sale
# transaction data
data("Groceries", package = "arules")
Groceries
```

To read a random CSV file:

```{r, eval=FALSE}
ds <- arules::read.transactions(
  file = "yet_another_flat_file.csv",
  format = "single",
  sep = ",",
  cols = c("Transaction ID", "Item Code"),
  rm.duplicates = T,
  header = T
)
```
