---
title: "ISLR-R: 4. Cross-Validation and the Bootstrap"
output:
  prettydoc::html_pretty:
    toc: true
    theme: cayman
    highlight: github
    df_print: paged
date: "2023-04-25"
editor_options:
  markdown:
    wrap: 79
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
import::from(magrittr, "%>%", "%$%", .into = "operators")
# import dataset
import::from(ISLR, Auto)
import::from(ISLR, Portfolio)
```

In this lab you will work through Section 5.3.

## The Validation Set Approach

The first thing to do is to load the necessary package and briefly explore the data set called **Auto**

```{r}
set.seed(10)
Auto %>% summary()
```

```{r}
names(Auto)
Auto[1:4, ]
```

```{r}
horsepower <- Auto$horsepower
mpg <- Auto$mpg
horselims <- range(horsepower)
horsepower.grid <- seq(from = horselims[1], to = horselims[2])
```

Define the training sample. Define the training sample

```{r}
train <- sample(392, 196)

# Fit first order polynomial
lm.fit <- lm(mpg ~ horsepower, data = Auto, subset = train)
mean((mpg - predict(lm.fit, Auto))[-train]^2)

# Fit a second order polynomial
lm.fit2 <- lm(mpg ~ poly(horsepower, 2), data = Auto, subset = train)
mean((mpg - predict(lm.fit2, Auto))[-train]^2)
preds2 <- predict(lm.fit2, newdata = list(horsepower = horsepower.grid))

# Fit a third order polynomial
lm.fit3 <- lm(mpg ~ poly(horsepower, 3), data = Auto, subset = train)
mean((mpg - predict(lm.fit3, Auto))[-train]^2)
preds3 <- predict(lm.fit3, newdata = list(horsepower = horsepower.grid))
```

Plotting:

```{r}
Auto %$% plot(horsepower, mpg)
points(horsepower[train], mpg[train], col = "red")
abline(lm.fit, col = "orange", lwd = 3)
lines(horsepower.grid, preds2, col = "green", lwd = 3)
lines(horsepower.grid, preds3, col = "purple", lwd = 3)
```

## k-Fold Cross-Validation

Here we are going to do the k-Fold Cross-Validation

- Use K-fold Cross-Validation (CV) when:
  - large dataset
  - model hard to train
- Use Leave-One-Out Cross-Validation (LOOCV) when:
  - small dataset
  - easy to train

```{r}
set.seed(1)
cv.error.10 <- rep(0, 10)
for (i in 1:10) {
  glm.fit <- glm(mpg ~ poly(horsepower, i), data = Auto)
  cv.error.10[i] <- boot::cv.glm(Auto, glm.fit, K = 10)$delta[1]
}
```

Calculate and plot the error:
```{r}
cv.error.10
plot(1:10, cv.error.10, type = "b", col = "red")
```

### One SE Rule

Prepare dataset:

```{r}
data <- Auto %>%
  dplyr::select("mpg", "horsepower") %>%
  dplyr::mutate(const = 1) %>%
  dplyr::mutate(horsepower2 = horsepower^2) %>%
  dplyr::mutate(horsepower3 = horsepower^3) %>%
  dplyr::mutate(horsepower4 = horsepower^4) %>%
  dplyr::mutate(horsepower5 = horsepower^5) %>%
  dplyr::mutate(horsepower6 = horsepower^6) %>%
  dplyr::mutate(horsepower7 = horsepower^7) %>%
  dplyr::mutate(horsepower8 = horsepower^8)
```

models

```{r}
control <- caret::trainControl(
  method = "cv",
  number = 10, savePredictions = "final"
)
get_model <- function(formula) {
  caret::train(formula,
    data = data, method = "lm", trControl = control
  )
}
mdl_00 <- get_model(mpg ~ const + 0)
mdl_01 <- get_model(mpg ~ horsepower + const + 0)
mdl_02 <- get_model(mpg ~ horsepower2 + horsepower + const + 0)
mdl_03 <- get_model(mpg ~ horsepower3 + horsepower2 + horsepower + const + 0)
mdl_04 <- get_model(
  mpg ~ horsepower4 + horsepower3 + horsepower2 + horsepower + const + 0
)
mdl_05 <- get_model(
  mpg ~ horsepower5 + horsepower4 + horsepower3 + horsepower2 + horsepower + const + 0
)
mdl_06 <- get_model(
  mpg ~ horsepower6 + horsepower5 + horsepower4 + horsepower3 + horsepower2 + horsepower + const + 0
)
mdl_07 <- get_model(
  mpg ~ horsepower7 + horsepower6 + horsepower5 + horsepower4 + horsepower3 + horsepower2 + horsepower + const + 0
)
mdl_08 <- get_model(
  mpg ~ horsepower8 + horsepower7 + horsepower6 + horsepower5 + horsepower4 + horsepower3 + horsepower2 + horsepower + const + 0
)
# Calculate the residuals for each fold
res_00 <- mdl_00$pred$obs - mdl_00$pred$pred
res_01 <- mdl_01$pred$obs - mdl_01$pred$pred
res_02 <- mdl_02$pred$obs - mdl_02$pred$pred
res_03 <- mdl_03$pred$obs - mdl_03$pred$pred
res_04 <- mdl_04$pred$obs - mdl_04$pred$pred
res_05 <- mdl_05$pred$obs - mdl_05$pred$pred
res_06 <- mdl_06$pred$obs - mdl_06$pred$pred
res_07 <- mdl_07$pred$obs - mdl_07$pred$pred
res_08 <- mdl_08$pred$obs - mdl_08$pred$pred
# Calculate the MSE for each fold
mse_00 <- tapply(res_00^2, mdl_00$pred$Resample, mean)
mse_01 <- tapply(res_01^2, mdl_01$pred$Resample, mean)
mse_02 <- tapply(res_02^2, mdl_02$pred$Resample, mean)
mse_03 <- tapply(res_03^2, mdl_03$pred$Resample, mean)
mse_04 <- tapply(res_04^2, mdl_04$pred$Resample, mean)
mse_05 <- tapply(res_05^2, mdl_05$pred$Resample, mean)
mse_06 <- tapply(res_06^2, mdl_06$pred$Resample, mean)
mse_07 <- tapply(res_07^2, mdl_07$pred$Resample, mean)
mse_08 <- tapply(res_08^2, mdl_08$pred$Resample, mean)

ds_mse <- data.frame(mse_00, mse_01, mse_02, mse_03, mse_04, mse_05, mse_06, mse_07, mse_08)
boxplot(ds_mse)
```

### Yet Another Model Selection (L2-MSE)

```{r}
get_indices <- function(ratio, n_rec) {
  n_sample <- round(ratio * n_rec, 0)
  return(sample(n_rec, n_sample))
}

N_REC <- data %>% nrow()

get_x <- function(ds, predictor, response) {
  model.matrix(eval(parse(text = paste(response, "~", predictor))),
    data = ds,
  )
}

get_y <- function(ds) {
  ds$mpg
}

get_beta <- function(x, y) {
  solve(t(x) %*% x) %*% (t(x) %*% y)
}

idx_test <- get_indices(0.25, N_REC)
ds_tr <- data[-idx_test, ]
ds_te <- data[idx_test, ]

y_tr <- get_y(ds_tr)
x1_tr <- get_x(ds_tr, "horsepower + const + 0", "mpg")
x2_tr <- get_x(ds_tr, "horsepower2 + horsepower + const + 0", "mpg")
x3_tr <- get_x(
  ds_tr,
  "horsepower3 + horsepower2 + horsepower + const + 0", "mpg"
)
x4_tr <- get_x(
  ds_tr,
  "horsepower4 + horsepower3 + horsepower2 + horsepower + const + 0", "mpg"
)

y_te <- get_y(ds_te)
x1_te <- get_x(ds_te, "horsepower + const + 0", "mpg")
x2_te <- get_x(ds_te, "horsepower2 + horsepower + const + 0", "mpg")
x3_te <- get_x(
  ds_te,
  "horsepower3 + horsepower2 + horsepower + const + 0", "mpg"
)
x4_te <- get_x(
  ds_te,
  "horsepower4 + horsepower3 + horsepower2 + horsepower + const + 0", "mpg"
)

beta1 <- get_beta(x1_tr, y_tr)
beta2 <- get_beta(x2_tr, y_tr)
beta3 <- get_beta(x3_tr, y_tr)

get_l2_mse <- function(x, y, beta, lambda = 0.001) {
  mean((y - x %*% beta)^2) + lambda * (t(beta) %*% beta)[1]
}

l2mse_0 <- mean((y_te - mean(y_te))^2)
l2mse_1 <- get_l2_mse(x1_te, y_te, beta1)
l2mse_2 <- get_l2_mse(x2_te, y_te, beta2)
l2mse_3 <- get_l2_mse(x3_te, y_te, beta3)

cat(l2mse_0, l2mse_1, l2mse_2, l2mse_3)
```

## The Bootstrap 

Let's bootstrap some "data set".

A function to calculate the optimal value (minimum variance)
```{r}
alpha <- function(data) {
  X <- data$X
  Y <- data$Y
  return((var(Y) - cov(X, Y)) / (var(X) + var(Y) - 2 * cov(X, Y)))
}
alpha(Portfolio)
```
What is the standard error of alpha?
```{r}
alpha.fn <- function(data, index) {
  X <- data$X[index]
  Y <- data$Y[index]
  return((var(Y) - cov(X, Y)) / (var(X) + var(Y) - 2 * cov(X, Y)))
}
```
  
check the function

```{r}
alpha.fn(Portfolio, 1:100)
```

one iteration

```{r}
set.seed(1)
alpha.fn(Portfolio, sample(100, 100, replace = T))
```

the bootstrap

```{r}
boot.out <- boot::boot(Portfolio, alpha.fn, R = 1000)
boot.out
plot(boot.out)
```

## Estimating the accuracy of a linear regression model

Recall the usual linear model estimation:

```{r}
lm(mpg ~ horsepower, data = Auto) %>% summary()
```

For the Auto data set, we want to use bootstrap to estimate the accuracy of the
linear regression model, without knowing the T-test formula:

```{r}
N_ROW <- Auto %>% nrow()
lm.fn <- function(data, index) {
  x <- data %$% horsepower[index]
  y <- data %$% mpg[index]
  b1 <- cov(x, y) / var(x)
  b0 <- mean(y) - b1 * mean(x)
  coef <- c(b0, b1)
  names(coef) <- c("interception", "slope")
  return(coef)
}
set.seed(1)
```

Estimate:

```{r}
lm.fn(Auto, sample(N_ROW, N_ROW, replace = F))
```

One bootstrap iteration:

```{r}
lm.fn(Auto, sample(N_ROW, N_ROW, replace = T))
```

Perform bootstrapping, and summarize the result

```{r}
res <- boot::boot(Auto, lm.fn, 100)
res
```
